### 2015-05-10 paper.tex
Created paper skeleton

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    PAPER for MA 952/672
%    Tony Zhang
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}

\usepackage{tony}

\title{Voice Classification by Neural Networks}
\author{Tony Zhang}

\begin{document}
\maketitle

\begin{abstract}
% abstract
\end{abstract}

\section{Introduction}

% introduce things
% theory background

\section{Methodology}

% number of data points
% feature vectors (and which ultimately chosen)
% structure of neural net
% gradient descent

\section{Results}

% ROC AUC scores and such

\end{document}

### 2015-05-14 paper.tex
Style file updated; significant portions of intro written

@@
 \usepackage{tony}
+
+\usepackage{tikz}
+\usetikzlibrary{matrix,chains,positioning,decorations.pathreplacing,arrows}
+
+\DeclareMathOperator{\softmax}{softmax}
 
@@
 
+Though still a relatively new field, machine learning, the study of algorithms that can ``learn" dynamically, holds much potential. 
+% blah blah blah
+
+A curious development in the field was the \emph{neural network}, a biologically inspired machine learning model.
+Indeed, we see how the nodes in \cref{fig:neuralnet} resemble neurons sending information to other neurons.
+We shall soon make these concepts mathematically rigorous.
+
+The goal of a neural network is to classify a \emph{feature vector}; that is, given some input data (``features", e.g.
+the batting average of a particular baseball player, the temperature that particular day), we would like to place this data point into one of a number of \emph{classes} (e.g.
+whether his/her next swing will/will not be a strike).
+In particular, the neural network will be a function taking input data and outputting some numbers that can be interpreted as the probability that this data point is in a particular class.\footnote{To be nitpicky, we should only say that these numbers are indices for the model's confidence that the data point is in some class.}
+
+Let's make that concrete.
+For our baseball example, suppose we'd like to predict whether Lauren the baseball player (batting average 0.315) will hit the ball on her next swing.
+The current temperature at the stadium is 21.6 degrees Celsius.
+So we give the neural net the feature vector $(0.315, 21.6)$.
+The neural net then outputs us $(0.21, 0.79)$---a strike with ``probability" 0.21 and a non-strike with ``probability" 0.79.
+
+% picture of neural net
+\begin{figure}[htbp]
+\centering
+\input{aux/neuralnet.tex}
+\caption{A simple neural network.}
+\label{fig:neuralnet}
+\end{figure}
+
+\subsection{Neural Net Theory}
+
+How does a neural network make these predictions?
+Each neural network takes the input and sends it through a set of \emph{neurons} (each carrying a value) organized into \emph{layers}.
+(\cref{fig:neuralnet} displays 3 layers, but a neural net can have arbitrarily many layers, with arbitrarily many neurons in each.)
+The value of each neuron is either predetermined (as are the ``bias" neurons in \cref{fig:neuralnet}, whose values are fixed at 1, or the input neurons, whose values are just the individual components of the feature vectors---one input neuron could contain a baseball player's batting average, for example) or computed from the values of neurons in the previous layer.
+
+For our purposes, each neuron $z$ (except for the bias neuron) in a \emph{hidden layer} (layer that isn't the first or last) depends on the neurons $x_i$ in the previous layer as follows:
+\[ z = \sigma\left(\sum_i \alpha_i x_i \right) \]
+where $\sigma(t) = (1 + \exp(-t))^{-1}$ is the logistic function and the $\alpha_i$ are some known constants (imagine these as unique ``weights" attached to each arrow between neurons in \cref{fig:neuralnet}).
+We shall see later where the values of $\alpha$ come from.
+
+The values of the neurons $y_k$ ($k = 1, \dots, K$) in the output layer, however, are defined as follows:
+\[ y_k = \softmax_k(n_1, n_2, \dots, n_K) = \frac{\exp(n_k)}{\sum_j \exp(n_j)} \]
+where each $n_k$ is a ``proto-value" associated with neuron $y_k$ with value
+\[ n_k = \sum_i \beta_i x_i \]
+where the $\beta_i$ are known constants (like the $\alpha_i$ from a little before) and the $x_i$ are the neurons in the penultimate layer.
+
+In general, with the exception of the input neurons and the bias neurons, each neuron's value is simply produced by applying some nonlinear function (e.g.
+$\sigma, \softmax$) to a linear combination of the values of the neurons in the previous layer.
+The weights on the linear combination are predetermined.
+Clearly, to teach a neural net, we must figure out the right weights for the situation.
+
+\subsection{Training}
+
 \section{Methodology}

### 2015-05-15 paper.tex
Added batch gradient descent section, updated tony.sty

@@
 
+% for drawing figures and stuff
 \usepackage{tikz}
@@
 
+% new math operator
 \DeclareMathOperator{\softmax}{softmax}
@@
 \begin{abstract}
-% abstract
+% TODO
+% neural net to classify voice gender (age?)
+% 
 \end{abstract}
@@
 Though still a relatively new field, machine learning, the study of algorithms that can ``learn" dynamically, holds much potential.
+Already, machine learning algorithms are finding great use in everything from identification of Higgs Boson decay events in high-energy particle collisions to recommendation systems on such sites as Facebook, YouTube, and Netflix.
 % blah blah blah
 
-A curious development in the field was the \emph{neural network}, a biologically inspired machine learning model.
+A curious development in macine learning was the \emph{neural network}, a biologically inspired machine learning model.
 Indeed, we see how the nodes in \cref{fig:neuralnet} resemble neurons sending information to other neurons.
-We shall soon make these concepts mathematically rigorous.
+We shall soon offer a mathematically rigorous definition of the neural network.
 
@@
 the batting average of a particular baseball player, the temperature that particular day), we would like to place this data point into one of a number of \emph{classes} (e.g.
-whether his/her next swing will/will not be a strike).
-In particular, the neural network will be a function taking input data and outputting some numbers that can be interpreted as the probability that this data point is in a particular class.\footnote{To be nitpicky, we should only say that these numbers are indices for the model's confidence that the data point is in some class.}
+this player's next swing will/will not be a strike).
+In particular, the neural network will be a function taking input data and outputting some numbers that can be interpreted as the probability that this data point is in a particular class.\footnote{To be nitpicky, we should only say that these numbers are simply indices for the model's confidence that the data point is in some class.}
 
@@
 
+% describe our project
+
 \subsection{Neural Net Theory}
@@
 For our purposes, each neuron $z$ (except for the bias neuron) in a \emph{hidden layer} (layer that isn't the first or last) depends on the neurons $x_i$ in the previous layer as follows:
-\[ z = \sigma\left(\sum_i \alpha_i x_i \right) \]
+\beq
+\label{eq:hiddenlayer}
+z = \sigma\left(\sum_i \alpha_i x_i \right)
+\eeq
 where $\sigma(t) = (1 + \exp(-t))^{-1}$ is the logistic function and the $\alpha_i$ are some known constants (imagine these as unique ``weights" attached to each arrow between neurons in \cref{fig:neuralnet}).
@@
 The values of the neurons $y_k$ ($k = 1, \dots, K$) in the output layer, however, are defined as follows:
-\[ y_k = \softmax_k(n_1, n_2, \dots, n_K) = \frac{\exp(n_k)}{\sum_j \exp(n_j)} \]
+\beq
+y_k = \softmax_k(n_1, n_2, \dots, n_K) = \frac{\exp(n_k)}{\sum_j \exp(n_j)}
+\eeq
 where each $n_k$ is a ``proto-value" associated with neuron $y_k$ with value
-\[ n_k = \sum_i \beta_i x_i \]
+\beq
+\label{eq:netvalue}
+n_k = \sum_i \beta_i x_i
+\eeq
 where the $\beta_i$ are known constants (like the $\alpha_i$ from a little before) and the $x_i$ are the neurons in the penultimate layer.
@@
 The weights on the linear combination are predetermined.
-Clearly, to teach a neural net, we must figure out the right weights for the situation.
+Clearly, ``teaching" a neural net reduces to determining  the correct weights for the situation.
 
-\subsection{Training}
+\subsection{Gradient Descent}
+
+% insert stuff about gradient descent!
+Formally, we can rephrase our problem as follows: Given a set of $K$ data points $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (for some number of parameters $w_i$, which we write as a vector), find the parameters such that we minimize some \emph{objective function} dependent on both data points and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
+For our purposes, we will minimize mean-square error, defined as
+\beq
+\label{eq:meansqerror}
+E(\vec{w}) = \sum_k E_k(\vec{w}) = \sum_k (y_k - f(\vec{w})(x_k))^2
+\eeq
+
+In practice, there will be numerous parameters $w_i$ (each $\alpha$ and $\beta$ in \cref{eq:hiddenlayer,eq:netvalue} is a $w_i$), so finding a global minimum will be challenging: setting the derivatives of $E$ to 0 and solving the resulting nonlinear equations (our $f$ will involve the nonlinear logistic $\sigma$ and softmax functions) is unfeasible.
+Instead, we will employ a technique known as gradient descent.
+
+The basic idea is simple.
+Suppose we are looking for low points in some terrain.
+We pick an arbitrary location to begin.
+Then, we walk downhill (in the direction with the steepest drop) a certain distance.
+We stop, then check which way will take us down the steepest path from our new location.
+We walk downhill a certain distance again, then repeat the process.
+Clearly, gradient descent will \emph{not} necessarily find the global minimum, but rather a local one.
+
+We formalize this idea as follows.
+Given an objective function $E$, we start at some arbitrary $\vec{w}_0$.
+To find a better estimate for $\vec{w}$, we iteratively go ``downhill" from $\vec{w}_{n-1}$ to $\vec{w}_{n}$, computing
+\beq
+\vec{w}_n = \vec{w}_{n-1} - \eta\vec{\nabla} E(\vec{w}_{n-1})
+\eeq
+where $\eta > 0$ is some fixed step size (the ``certain distance" in ou
+r hill analogy).
+The gradient $\vec{\nabla} E$ represents the direction of steepest ascent of the function $E$.
+Gradient descent is extremely useful for our purposes because for the $E$ we are using, the resulting derivatives $\frac{\partial E}{\partial w_i}$ are relatively simple to determine.
+
+Still, since our objective function sums over all $K$ our data points $(x_k, y_k)$ (see \cref{eq:meansqerror}), gradient descent quickly becomes computationally expensive for us as we try to compute $O(K)$ partial derivatives.
+Is there a more efficient alternative?
+
+% TODO stochastic gradient descent
+
+\subsection{Receiver Operating Characteristic}
+
+% TODO
+% ROC curves
+% AUC metric
+
+\subsection{Fourier Analysis}
+
+% TODO
+% discrete fourier analysis
 
@@
 
+% TODO
 % number of data points
-% feature vectors (and which ultimately chosen)
+% feature vectors (and which ultimately chosen): Tony
 % structure of neural net
 % gradient descent
+
+% We collected XXXXXXXX voice samples, of which YYYYY were male and ZZZZZ were female.
 

### 2015-05-16 paper.tex
Started stochastic gradient descendescent & ROC sections; minor edits

@@
 
-\documentclass{article}
+\documentclass[10pt]{article}
 
@@
 
-A curious development in macine learning was the \emph{neural network}, a biologically inspired machine learning model.
+A curious development in macine learning was the \emph{neural network} (or informally, a neural net), a biologically inspired machine learning model.
 Indeed, we see how the nodes in \cref{fig:neuralnet} resemble neurons sending information to other neurons.
@@
 % insert stuff about gradient descent!
-Formally, we can rephrase our problem as follows: Given a set of $K$ data points $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (for some number of parameters $w_i$, which we write as a vector), find the parameters such that we minimize some \emph{objective function} dependent on both data points and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
+Formally, we can rephrase our problem as follows: Given a set of $K$ data points $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (where we write the parameters as components of vector $\vec{w}$), find the parameters that minimize some \emph{objective function} dependent on both the data points and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
 For our purposes, we will minimize mean-square error, defined as
@@
 
-In practice, there will be numerous parameters $w_i$ (each $\alpha$ and $\beta$ in \cref{eq:hiddenlayer,eq:netvalue} is a $w_i$), so finding a global minimum will be challenging: setting the derivatives of $E$ to 0 and solving the resulting nonlinear equations (our $f$ will involve the nonlinear logistic $\sigma$ and softmax functions) is unfeasible.
+In practice, there will be numerous parameters (each $\alpha$ and $\beta$ in \cref{eq:hiddenlayer,eq:netvalue} is a component of $\vec{w}$), so finding a global minimum will be challenging: setting the derivatives of $E$ to 0 and solving the resulting nonlinear equations (our $f$ will involve the nonlinear logistic $\sigma$ and softmax functions) is unfeasible.
 Instead, we will employ a technique known as gradient descent.
@@
 We pick an arbitrary location to begin.
-Then, we walk downhill (in the direction with the steepest drop) a certain distance.
+Then, we walk downhill (in the direction of steepest descent) a certain distance.
 We stop, then check which way will take us down the steepest path from our new location.
-We walk downhill a certain distance again, then repeat the process.
+We walk downhill in that direction a certain distance again, then repeat the process.
 Clearly, gradient descent will \emph{not} necessarily find the global minimum, but rather a local one.
@@
 We formalize this idea as follows.
-Given an objective function $E$, we start at some arbitrary $\vec{w}_0$.
+Given an objective function $E$, we start with some arbitrary set of weights $\vec{w}_0$.
 To find a better estimate for $\vec{w}$, we iteratively go ``downhill" from $\vec{w}_{n-1}$ to $\vec{w}_{n}$, computing
 \beq
+\label{eq:graddescent}
 \vec{w}_n = \vec{w}_{n-1} - \eta\vec{\nabla} E(\vec{w}_{n-1})
 \eeq
-where $\eta > 0$ is some fixed step size (the ``certain distance" in ou
-r hill analogy).
+where $\eta > 0$ is some fixed step size (the ``certain distance" in our hill analogy).
 The gradient $\vec{\nabla} E$ represents the direction of steepest ascent of the function $E$.
-Gradient descent is extremely useful for our purposes because for the $E$ we are using, the resulting derivatives $\frac{\partial E}{\partial w_i}$ are relatively simple to determine.
+Gradient descent is extremely useful for our purposes because for the $E$ we are using, as the resulting derivatives $\frac{\partial E}{\partial w}$ are relatively simple to determine.
 
@@
 
-% TODO stochastic gradient descent
+\subsubsection{Stochastic Gradient Descent}
+
+Returning to \cref{eq:meansqerror}, we find that our error can be expressed as the sum of the errors on each individual data point.
+Instead of minimizing $E = \sum_k E_k$, we can try to minimize an \emph{individual} $E_k$ (imagine replacing $E$ with $E_k$ in \cref{eq:graddescent}).
+In the long run, we have reason to expect this procedure also minimizes $E$.
+
+% TODO continue
 
@@
 
-% TODO
-% ROC curves
-% AUC metric
+Though our goal will ultimately be to find the feature vectors and training procedure that will result in the \emph{best} model, we have not yet introduced any metric for evaluating our models.
+We shall now do just that.
+
+% ROC space (false pos, true pos)
+% ROC curve
+% AUC
 

### 2015-05-17 paper.tex
Separated sentences into individual lines for future git-diff convenience
Addded to ROC section

@@
 
-The goal of a neural network is to classify a \emph{feature vector}; that is, given some input data (``features", e.g.
-the batting average of a particular baseball player, the temperature that particular day), we would like to place this data point into one of a number of \emph{classes} (e.g.
+The goal of a neural network is to classify \emph{instances} described by \emph{feature vectors}; that is, given some data about an instance (``features", e.g.
+the batting average of a particular baseball player, the temperature that particular day), we would like to place the instance into one of a number of \emph{classes} (e.g.
 this player's next swing will/will not be a strike).
-In particular, the neural network will be a function taking input data and outputting some numbers that can be interpreted as the probability that this data point is in a particular class.\footnote{To be nitpicky, we should only say that these numbers are simply indices for the model's confidence that the data point is in some class.}
+In particular, the neural network will be a function taking input data and outputting some numbers that can be interpreted as the probability that this instance is in a particular class.\footnote{To be nitpicky, we should only say that these numbers are simply indices for the model's confidence that the instance is in some class.}
 
@@
 
-How does a neural network make these predictions?
-Each neural network takes the input and sends it through a set of \emph{neurons} (each carrying a value) organized into \emph{layers}.
+What exactly is a neural net, and how does it make these predictions?
+Each neural net takes some input and sends it through a set of \emph{neurons}, each carrying a value, organized into \emph{layers}.
 (\cref{fig:neuralnet} displays 3 layers, but a neural net can have arbitrarily many layers, with arbitrarily many neurons in each.)
 The value of each neuron is either predetermined (as are the ``bias" neurons in \cref{fig:neuralnet}, whose values are fixed at 1, or the input neurons, whose values are just the individual components of the feature vectors---one input neuron could contain a baseball player's batting average, for example) or computed from the values of neurons in the previous layer.
+We read off the output of the neural net in the final layer, each of whose neurons will give us values in $(0, 1)$.
+There is one neuron per class, each representing the level of certainty the model has that our instance belongs to the respective class of the neuron.
 
@@
 The weights on the linear combination are predetermined.
-Clearly, ``teaching" a neural net reduces to determining  the correct weights for the situation.
+Clearly, ``training" a neural net reduces to determining the correct weights for the situation.
 
@@
 % insert stuff about gradient descent!
-Formally, we can rephrase our problem as follows: Given a set of $K$ data points $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (where we write the parameters as components of vector $\vec{w}$), find the parameters that minimize some \emph{objective function} dependent on both the data points and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
+Formally, we can rephrase our problem as follows: Given a set of $K$ instances $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (where we write the parameters as components of vector $\vec{w}$), find the parameters that minimize some \emph{objective function} dependent on both the instances and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
 For our purposes, we will minimize mean-square error, defined as
@@
 
-Still, since our objective function sums over all $K$ our data points $(x_k, y_k)$ (see \cref{eq:meansqerror}), gradient descent quickly becomes computationally expensive for us as we try to compute $O(K)$ partial derivatives.
+Still, since our objective function sums over all $K$ our instances $(x_k, y_k)$ (see \cref{eq:meansqerror}), gradient descent quickly becomes computationally expensive for us as we try to compute $O(K)$ partial derivatives.
 Is there a more efficient alternative?
@@
 
-Returning to \cref{eq:meansqerror}, we find that our error can be expressed as the sum of the errors on each individual data point.
+Returning to \cref{eq:meansqerror}, we find that our error can be expressed as the sum of the errors on each individual instance.
 Instead of minimizing $E = \sum_k E_k$, we can try to minimize an \emph{individual} $E_k$ (imagine replacing $E$ with $E_k$ in \cref{eq:graddescent}).
@@
 
-\subsection{Receiver Operating Characteristic}
+\subsection{Receiver Operating Characteristic (ROC)}
 
@@
 
+For a binary classification model, where each instance is classified as in or not in some particular group (e.g.
+the next swing will/won't be a strike), we define the \emph{true positive rate} as the fraction of the instances in the group that are classified as such by the model.
+Similarly, the \emph{false positive rate} is the fraction of the instances \emph{not} in the group that are still classified as in the group by the model.
+A disease test with true and false positive rates 99\% and 5\% will successfully detect disease 99\% of the time.
+On the other hand, it will detect disease in 5\% of all uninfected individuals.
+Let us say that this disease test falls at $(0.05, 0.99)$ in \emph{receiver operating characteristic (ROC) space}, a coordinate plane where false and true positive rates are plotted on the $x$ and $y$ axes, respectively.
+
+Recall that our neural net has \emph{continuous output}---rather than tell us whether some instance is in a class or not, it gives us a number in $(0, 1)$ that serves as an index of its confidence that the instance is in the class.
+To convert this continuous output into a binary prediction (in/not in the class), we can set a threshold for this index.
+Those instances for which the neural net gives us a result above the threshold are considered predicted to be in the class, and \emph{vice versa}.
+
+As we vary the threshold, the neural net's position in ROC space will vary.
+As the threshold moves from 0 to 1, the neural net will trace out a curve in ROC space (see \cref{fig:rocexample}, which we will call the model's \emph{ROC curve}.
+% ROC curve intuition
+
+\begin{figure}
+\centering
+% \includegraphics{aux/rocexample.pdf}
+\caption{An example ROC curve.}
+\label{fig:rocexample}
+\end{figure}
+% TODO include
+
 % ROC space (false pos, true pos)
@@
 % AUC
+% insert figure
 
@@
 % TODO
-% number of data points
+% number of instances
 % feature vectors (and which ultimately chosen): Tony

### 2015-05-18 paper.tex
ROC section basically finished

@@
 As we vary the threshold, the neural net's position in ROC space will vary.
-As the threshold moves from 0 to 1, the neural net will trace out a curve in ROC space (see \cref{fig:rocexample}, which we will call the model's \emph{ROC curve}.
-% ROC curve intuition
+In particular, as we vary threshold from 0 to 1, the neural net will trace out a curve in ROC space (see \cref{fig:rocexample}), which we will call the model's \emph{ROC curve}.
+Note that a random classifier, with no distinguishing power between positive and negative instances, would have an ROC curve coinciding with the line $y = x$, as we expect equal true and false positive rates from this classifier.
+On the other hand, a model whose ROC curve lies above this line would tend to give higher ``scores" (the continuous output) to positive instances than to negative ones.
+Thus, we would like our model's ROC curve to lie as far above the line $y = x$ as possible.
+
+This intuition suggests that a good metric for evaluating our models could be the area under the ROC curve.
+Indeed, \emph{Area Under Curve} (AUC) is a common metric for binary classification.
+Its value can range from 0 to 1, though in practice, our values will lie in $(0.5, 1)$.
+A value of 0.5 corresponds to a model no better than a random classifier, while a value of 1 corresponds to a model that perfectly distinguishes between positive and negative instances.
 
@@
 \centering
-% \includegraphics{aux/rocexample.pdf}
-\caption{An example ROC curve.}
+\includegraphics[width=\textwidth]{aux/rocexample.pdf}
+\caption{An example ROC curve, with AUC $\approx 0.7804$. The dashed line shows $y = x$.}
 \label{fig:rocexample}
 \end{figure}
-% TODO include
-
-% ROC space (false pos, true pos)
-% ROC curve
-% AUC
-% insert figure
 

### 2015-05-18 paper.tex
updated authors, started abstract, minor edits

@@
 \title{Voice Classification by Neural Networks}
-\author{Tony Zhang}
+\author{Tony Zhang, James Tsui, Maryam Bahrani}
 
@@
 %
+
+In this paper, we describe a machine learning schema for voice classification by sex. We collected %XXXXX
+voice samples that we used to train an artificial neural net with a single hidden layer using stochastic gradient descent.
+After experimenting to determine the optimal training parameters (number of iterations, etc.), using %%10-fold cross-validation with ROC AUC as our metric to evaluate our model.
+With % certain settings
+we were able to produce a model that performs %stats
 \end{abstract}
@@
 % insert stuff about gradient descent!
-Formally, we can rephrase our problem as follows: Given a set of $K$ instances $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (where we write the parameters as components of vector $\vec{w}$), find the parameters that minimize some \emph{objective function} dependent on both the instances and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
+Formally, we can rephrase our problem as follows: Given a set of $K$ instances $(x_k, y_k)$ and a family of functions $f(\vec{w})$,\footnote{Note that $f$ itself is a function $f:W\to X$, where $W$ is the space of all possible combinations of weights and $X$ is the space of functions from inputs---our features---to outputs---our predictions.
+Thus, $\vec{w}$ is a vector whose components are the parameters that determine our model.} find the parameters that minimize the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
+In particular, we are minimizing some \emph{objective function} dependent on both the instances and the parameters that quantifies that inaccuracy.
 For our purposes, we will minimize mean-square error, defined as
@@
 
+\subsection{Cross Validation}
+
+% why do we want it? +overfitting
+% training/test data
+%
+
 \subsection{Fourier Analysis}

### 2015-05-27 paper.tex
Methodology section completed up to Fourier

@@
 \title{Voice Classification by Neural Networks}
-\author{Tony Zhang, James Tsui, Maryam Bahrani}
+\author{
+	\href{mailto:txz@mit.edu}{Tony Zhang},
+	\href{mailto:jtsui@hotchkiss.org}{James Tsui},
+	\href{mailto:mbahrani@hotchkiss.org}{Maryam Bahrani}
+}
 
@@
 % training/test data
-%
 
@@
 
+Subjects were selected based on availability, with a majority of participants being students of the Hotchkiss School.
+Voice recordings were collected from a total of $n = 63$ subjects (27 male and 36 female).
+
+Subject voice data was obtained in the form of 22050 Hz \texttt{.aiff} files, recorded using the microphone of an iPhone.
+Subjects were instructed to speak until they were told to stop, and an approximately 10 second recording of their speech was made.
+The sex and age of the subjects were recorded.
+Although recorded, subject’s age is likely to be insignificant due to the stabilization of most voices post-puberty.
+Some subjects were given novels and other reading material to read from, as they were uncomfortable with speaking without directives.
+This data was collected in relatively silent locations, thus minimizing background or stray noises.
+
+All data processing and analysis was completed in Wolfram Mathematica.
+The voice samples were imported and processed using Fourier analysis.
+We experimented with multiple schemata for feature extraction, before ultimately settling on the best for our purposes.
+Each voice sample was broken into 0.5 second fragments (any trailing data at the end of the sound sample was discarded).
+We took the norms of the discrete Fourier transform (DFT) of each fragment and added them together (i.e.
+if fragments $1, 2, \dots, n$ have DFTs $\hat{f_1},\hat{f_2}, \dots, \hat{f_n}$, we find $g = \sum_k |\hat{f_k}|$).
+The resulting function $g$ was normalized as in \[\bar{g} = \frac{g}{\sum\limits_m g(m)}\] where we sum over all $m$ in the (finite) domain of $g$.\footnote{As we are working with half-second fragments of 22050 Hz sound files, this domain is actually $\frac{\ZZ}{11025\ZZ}$.} Each voice file was analyzed thus, and a series of features were generated for each; these features constituted a profile of each subject's voice.
+
 \section{Results}

### 2015-05-27 paper.tex
MAJOR EDITS OMG --- continued methodology

@@
 In the long run, we have reason to expect this procedure also minimizes $E$.
-
-% TODO continue
+Indeed, stochastic gradient descent is the industry standard for training artificial neural nets.
 
@@
 
-\subsection{Fourier Analysis}
-
-% TODO
-% discrete fourier analysis
+%\subsection{Fourier Analysis}
+%
+%% TODO
+%% discrete fourier analysis
 
@@
 % TODO
-% number of instances
-% feature vectors (and which ultimately chosen): Tony
 % structure of neural net
 % gradient descent
-
-% We collected XXXXXXXX voice samples, of which YYYYY were male and ZZZZZ were female.
 
@@
 The voice samples were imported and processed using Fourier analysis.
-We experimented with multiple schemata for feature extraction, before ultimately settling on the best for our purposes.
+We experimented with multiple schemata for feature extraction, before ultimately settling on the best for our purposes (as evaluated by the AUC metric), which we outline here.
 Each voice sample was broken into 0.5 second fragments (any trailing data at the end of the sound sample was discarded).
@@
 if fragments $1, 2, \dots, n$ have DFTs $\hat{f_1},\hat{f_2}, \dots, \hat{f_n}$, we find $g = \sum_k |\hat{f_k}|$).
-The resulting function $g$ was normalized as in \[\bar{g} = \frac{g}{\sum\limits_m g(m)}\] where we sum over all $m$ in the (finite) domain of $g$.\footnote{As we are working with half-second fragments of 22050 Hz sound files, this domain is actually $\frac{\ZZ}{11025\ZZ}$.} Each voice file was analyzed thus, and a series of features were generated for each; these features constituted a profile of each subject's voice.
+The resulting function $g$ was normalized as in \[\bar{g} = \frac{g}{\sum\limits_m g(m)}\] where we sum over all $m$ in the (finite) domain of $g$.
+(As we are working with half-second fragments of 22050 Hz sound files, this domain is actually $D = \frac{\ZZ}{11025\ZZ}$.)
+We then generated a series of features $x_k$ for each voice by finding the total value of $g$ within some predefined bins $B_i\subset D$.
+In our case, our bins were contiguous sequences of consecutive integers.\footnote{The smallest elements in each bin were: $\{25, 50, 75, 100, 125, 150, 175, 200, 250, 300, 350, 400, 600, 800, 1600\}$.
+Thus, our bins were 25 to 49, 50 to 74, etc.
+The final bin was 1600 to 3199.
+As we are working with 0.5 second fragments, these bins correspond to frequencies 50 to 98 Hz, 100 to 149 Hz, etc.}
+
+Each voice file was analyzed thus, and a series of features were generated for each; these features constituted a profile of each subject's voice.
+These features were then standardized so as to have mean 0 and standard deviation 1 (in effect, we replace them with $z$-scores).
+That is, if $\mu$ is the mean of feature $j$ across all voice samples and $\sigma$ is its standard deviation, the original value $x$ of the feature will be standardized to $z = \frac{x - \mu}{\sigma}$.
+
+These standardized feature vectors were used to train our neural net, to which we gave a single hidden layer with 8 nodes.
+To train our model, we turned to stochastic gradient descent, the standard for artificial neural networks.
+We experimented with the parameters involved in training our model (number of gradient descent iterations and step size $\eta$---see \cref{eq:graddescent}), finally settling on 1000 iterations and $\eta = 0.1$, which seemed to produce optimal results (see \cref{sec:results} for our evaluation scheme).
 
 \section{Results}
+\label{sec:results}
 

### 2015-05-27 paper.tex
paper basically done except for cross-validation

@@
 \begin{abstract}
-% TODO
-% neural net to classify voice gender (age?)
-%
+% TODO%
 
-In this paper, we describe a machine learning schema for voice classification by sex. We collected %XXXXX
-voice samples that we used to train an artificial neural net with a single hidden layer using stochastic gradient descent.
-After experimenting to determine the optimal training parameters (number of iterations, etc.), using %%10-fold cross-validation with ROC AUC as our metric to evaluate our model.
-With % certain settings
-we were able to produce a model that performs %stats
+In this paper, we describe a machine learning schema for voice classification by sex.
+We collected 63 voice samples (27 male, 36 female) that we used to train an artificial neural net with a single hidden layer using stochastic gradient descent.
+After experimenting to determine the optimal training parameters, evaluating each ensemble of parameters with repeated random sub-sampling validation with ROC AUC as our evaluation metric, we arrived at a set of parameters that yielded a score of $0.929 \pm 0.010$ (error given as std.
+err.
+of mean).
 \end{abstract}
@@
 \subsection{Receiver Operating Characteristic (ROC)}
+\label{subsec:roc}
 
@@
 Its value can range from 0 to 1, though in practice, our values will lie in $(0.5, 1)$.
-A value of 0.5 corresponds to a model no better than a random classifier, while a value of 1 corresponds to a model that perfectly distinguishes between positive and negative instances.
+A value of 0.5 corresponds to a model no better than a random classifier (which has ROC curve $y = x$, and therefore an AUC of 0.5), while a value of 1 corresponds to a model that perfectly distinguishes between positive and negative instances.
 
@@
 
-% ROC AUC scores and such
+To evaluate our model and to optimize our parameters, we used repeated random sub-sampling validation for each ensemble of parameters we tested.
+For each of the 30 evaluation rounds, we set aside roughly 30\% of the data for testing.
+Each round produced an AUC score (see \cref{subsec:roc}); the average AUC score over all 30 rounds was used as the final score for the ensemble of parameters.
+Our final parameters (8 hidden nodes, 1000 stochastic gradient descent iterations, $\eta = 0.1$) yielded an average AUC score of $0.929 \pm 0.010$.
+\footnote{Error is given as standard error of the mean of all $N = 30$ AUC scores from the evaluation rounds, computed as $\f{\sigma}{\sqrt{N}}$, where $\sigma$ is the standard deviation of the AUC scores.}
 

### 2015-05-27 paper.tex
first draft

@@
 \author{
-	\href{mailto:txz@mit.edu}{Tony Zhang},
-	\href{mailto:jtsui@hotchkiss.org}{James Tsui},
-	\href{mailto:mbahrani@hotchkiss.org}{Maryam Bahrani}
+	{Tony Zhang},
+	{James Tsui},
+	{Maryam Bahrani}
 }
@@
 Though still a relatively new field, machine learning, the study of algorithms that can ``learn" dynamically, holds much potential.
-Already, machine learning algorithms are finding great use in everything from identification of Higgs Boson decay events in high-energy particle collisions to recommendation systems on such sites as Facebook, YouTube, and Netflix.
-% blah blah blah
+Already, machine learning algorithms are finding great use in everything from identification of Higgs Boson decay events in high-energy particle collisions to recommender systems on such sites as Facebook, YouTube, and Netflix.
 
-A curious development in macine learning was the \emph{neural network} (or informally, a neural net), a biologically inspired machine learning model.
+A curious development in machine learning was the \emph{neural network} (or informally, a neural net), a biologically inspired machine learning model.
 Indeed, we see how the nodes in \cref{fig:neuralnet} resemble neurons sending information to other neurons.
@@
 
-The goal of a neural network is to classify \emph{instances} described by \emph{feature vectors}; that is, given some data about an instance (``features", e.g.
+The goal of a neural network is to detect patterns in datasets.
+In this project, we hoped to use these patterns to classify \emph{instances} described by \emph{feature vectors}; that is, given some data about an instance (``features", e.g.
 the batting average of a particular baseball player, the temperature that particular day), we would like to place the instance into one of a number of \emph{classes} (e.g.
@@
 Let's make that concrete.
-For our baseball example, suppose we'd like to predict whether Lauren the baseball player (batting average 0.315) will hit the ball on her next swing.
+For our baseball example, suppose we would like to predict whether Lauren the baseball player (batting average 0.315) will hit the ball on her next swing.
 The current temperature at the stadium is 21.6 degrees Celsius.
@@
 where the $\beta_i$ are known constants (like the $\alpha_i$ from a little before) and the $x_i$ are the neurons in the penultimate layer.
+Note that by definition of the softmax function, $\sum_k y_k = 1$.
 
@@
 % insert stuff about gradient descent!
-Formally, we can rephrase our problem as follows: Given a set of $K$ instances $(x_k, y_k)$ and a family of functions $f(\vec{w})$,\footnote{Note that $f$ itself is a function $f:W\to X$, where $W$ is the space of all possible combinations of weights and $X$ is the space of functions from inputs---our features---to outputs---our predictions.
-Thus, $\vec{w}$ is a vector whose components are the parameters that determine our model.} find the parameters that minimize the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
+Formally, we can rephrase our problem as follows: Given a set of $K$ instances $(x_k, y_k)$ and a family of functions $f(\vec{w})$,\footnote{Note that $f$ itself is a function $f:W\to X$, where $W$ is the space of all possible combinations of weights and $X$ is the space of functions from inputs (our features) to outputs (our predictions).
+Thus, $\vec{w}$ is a vector whose components are the parameters that determine our model.} find the parameters (components of $\vec{w}$) that minimize the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$.
 In particular, we are minimizing some \emph{objective function} dependent on both the instances and the parameters that quantifies that inaccuracy.
@@
 Given an objective function $E$, we start with some arbitrary set of weights $\vec{w}_0$.
+In practice, we start with randomly generated weights close to 0.
 To find a better estimate for $\vec{w}$, we iteratively go ``downhill" from $\vec{w}_{n-1}$ to $\vec{w}_{n}$, computing
@@
 The gradient $\vec{\nabla} E$ represents the direction of steepest ascent of the function $E$.
-Gradient descent is extremely useful for our purposes because for the $E$ we are using, as the resulting derivatives $\frac{\partial E}{\partial w}$ are relatively simple to determine.
+Gradient descent is extremely useful for the $E$ we are using, as the resulting derivatives $\frac{\partial E}{\partial w}$ are relatively simple to determine.
 
-Still, since our objective function sums over all $K$ our instances $(x_k, y_k)$ (see \cref{eq:meansqerror}), gradient descent quickly becomes computationally expensive for us as we try to compute $O(K)$ partial derivatives.
+Still, since our objective function sums over all $K$ of our instances $(x_k, y_k)$ (see \cref{eq:meansqerror}), gradient descent quickly becomes computationally expensive for us as we try to compute $O(K)$ partial derivatives.
 Is there a more efficient alternative?
@@
 For a binary classification model, where each instance is classified as in or not in some particular group (e.g.
-the next swing will/won't be a strike), we define the \emph{true positive rate} as the fraction of the instances in the group that are classified as such by the model.
+the next swing will/will not be a strike), we define the \emph{true positive rate} as the fraction of the instances \emph{in} the group that are classified as such by the model.
 Similarly, the \emph{false positive rate} is the fraction of the instances \emph{not} in the group that are still classified as in the group by the model.
-A disease test with true and false positive rates 99\% and 5\% will successfully detect disease 99\% of the time.
+For example, a disease test with true and false positive rates 99\% and 5\% will successfully detect disease 99\% of the time.
 On the other hand, it will detect disease in 5\% of all uninfected individuals.
@@
 
-Recall that our neural net has \emph{continuous output}---rather than tell us whether some instance is in a class or not, it gives us a number in $(0, 1)$ that serves as an index of its confidence that the instance is in the class.
+Recall that our neural net has \emph{continuous output}---instead of telling us whether some instance is in a class or not, it gives us a number in $(0, 1)$ that serves as an index of its confidence that the instance is in the class.
 To convert this continuous output into a binary prediction (in/not in the class), we can set a threshold for this index.
@@
 Note that a random classifier, with no distinguishing power between positive and negative instances, would have an ROC curve coinciding with the line $y = x$, as we expect equal true and false positive rates from this classifier.
-On the other hand, a model whose ROC curve lies above this line would tend to give higher ``scores" (the continuous output) to positive instances than to negative ones.
+On the other hand, a model whose ROC curve lies above this line would tend to give higher ``scores" (the output of our neural net prior to thresholding) to positive instances than to negative ones.
 Thus, we would like our model's ROC curve to lie as far above the line $y = x$ as possible.
@@
 
-% why do we want it? +overfitting
-% training/test data
+We want the neural net we generate to be robust and generalizable.
+A good model should sufficiently capture any signal in the training set.
+If the model, on the other hand, fails to detect some significant signal, we call our model \emph{underfitted}.
+Additionally, we do not want the model to track the training set too closely, treating its random noise as meaningful correlations, thus \emph{overfitting} the data by failing to generalize to unseen data.
+How do we evaluate a model in a manner that discourages underfitting and overfitting?
+
+Cross-validation is a statistical technique used to aid in model selection that addresses these issues.
+While there is great variation in the methods of cross-validation, the procedure generally consists of separating the data into a training set and testing set.
+The training data is used to train the model, and the testing data (and only the testing data) is used for the evaluation of the model, thus discouraging overfitting to the training data.
+For our purposes, we decided to cross-validate using the repeated random sub-sampling method.
+This method randomly allocates a certain percentage of the data for training (we use 70\%, a standard amount), and the rest for testing.
+Repeated random subsampling reruns this process for a fixed nubmer of iterations, keeping track of some performance metric each round.
+These performance measurements are then aggregated to give an estimate of how the model performs on new data.
 
@@
 The sex and age of the subjects were recorded.
-Although recorded, subject’s age is likely to be insignificant due to the stabilization of most voices post-puberty.
+Although recorded, subjects' age is likely to be insignificant due to the stabilization of most voices post-puberty.
 Some subjects were given novels and other reading material to read from, as they were uncomfortable with speaking without directives.
